Another paper with lots of text and fancy pictures

### 2023-12-07 paper.tex
More about shift operator

\section{Computational algorithms}
Three-point correlation function $S_3$ can be computed either in spatial domain
or in frequency domain via the convolution theorem. The simplest algorithm below
calculates $S_3$ pointwise in spatial domain:
\begin{algorithmic}[1]
  \Procedure{$S_3$}{$A, n, x_1, x_2$}
  \State $A_n \gets I_n (A)$
  \State $A'_n \gets \mathfrak{S}(A_n, x_1)$
  \State $A''_n \gets \mathfrak{S}(A_n, x_2)$
  \State $T \gets A_n \cdot A'_n \cdot A''_n$
  \State \textbf{return} $Sum(T) / Norm(A, x_1, x_2)$
  \EndProcedure
\end{algorithmic}
In this algorithm the indicator function $I_n$ for the phase of interest $n$ is
applied to the input array $A$. Then the shift operator $\mathfrak{S}$ is used
to shift $A_n$ by vectors $x_1$ and $x_2$. Finally, the three arrays are
multiplied pointwise and normalized sum of all elements of the product is
returned.

There are two common boundary conditions when applying $\mathfrak{S}$. The first
condition is to extend $A_n$ periodically (\cref{fig:s3-periodic}) when
accessing out-of-bounds array elements. This is the same as circular-shifting
$A_n$. In this case $Norm(A, x_1, x_2)$ does not depend on $x_1$ and $x_2$ and
is equal to the total number of elements in $A$. The second condition is to
replace out-of-bound elements with zeros (so-called zero padding)
(\cref{fig:s3-zeros}). A formula for $Norm(A, x_1, x_2)$ in this case is in
\ref{sec:number-of-trials}.
\begin{figure*}[tp]
  \centering
  \subfigure[Periodic boundary conditions]{
    \includegraphics[width=0.45\linewidth]{images/periodic.png}
    \label{fig:s3-periodic}}
  \hfill
  \subfigure[Zero padding]{
    \includegraphics[width=0.45\linewidth]{images/zeros.png}
    \label{fig:s3-zeros}}
  \caption[]{Computation of unnormalized $S_3$ function at point $(1, 2)$ for
    one-dimensional sequence of length 8.}
  \label{fig:s3-computation}
\end{figure*}

@@

Algorithms for three-point surface functions are the same as for $S_3$,
containing only small modifications. The main difference is that we need to
replace an indicator function $I_n$ with an edge detection operator $E_n$
(see \ref{sec:filter}). Then the algorithms are straightforward:
\begin{algorithmic}[1]
  \Procedure{$F_{sss}$}{$A, n, x_1, x_2$}
  \State $A_n \gets E_n (A)$
  \State $A'_n \gets \mathfrak{S}(A_n, x_1)$
  \State $A''_n \gets \mathfrak{S}(A_n, x_2)$
  \State $T \gets A_n \cdot A'_n \cdot A''_n$
  \State \textbf{return} $Sum(T) / Norm(A, x_1, x_2)$
  \EndProcedure

  \Procedure{$F_{ssv}$}{$A, n, x_1, x_2$}
  \State $A_n \gets E_n (A)$
  \State $A_{(void)} \gets I_{(void)} (A)$
  \State $A'_n \gets \mathfrak{S}(A_n, x_1)$
  \State $A'_{(void)} \gets \mathfrak{S}(A_{(void)}, x_2)$
  \State $T \gets A_n \cdot A'_n \cdot A'_{(void)}$
  \State \textbf{return} $Sum(T) / Norm(A, x_1, x_2)$
  \EndProcedure

  \Procedure{$F_{svv}$}{$A, n, x_1, x_2$}
  \State $A_n \gets E_n (A)$
  \State $A_{(void)} \gets I_{(void)} (A)$
  \State $A'_{(void)} \gets \mathfrak{S}(A_{(void)}, x_1)$
  \State $A''_{(void)} \gets \mathfrak{S}(A_{(void)}, x_2)$
  \State $T \gets A_n \cdot A'_{(void)} \cdot A''_{(void)}$
  \State \textbf{return} $Sum(T) / Norm(A, x_1, x_2)$
  \EndProcedure
\end{algorithmic}

@@

To compute $C_3$ function we need to use a connected-component labeling
algorithm \cite{4728561,PhysRevB.14.3438} to label clusters. Then, again, a
slightly modified algorithm for $S_3$ is used.
\begin{algorithmic}[1]
  \Procedure{$C_3$}{$A, n, x_1, x_2$}
  \State $A_n \gets I_n (A)$
  \State $C_n \gets \mathfrak{C}(A_n)$
  \State $C'_n \gets \mathfrak{S}(C_n, x_1)$
  \State $C''_n \gets \mathfrak{S}(C_n, x_2)$
  \State $T \gets C_n \odot C'_n \odot C''_n$
  \State \textbf{return} $Sum(T) / Norm(A, x_1, x_2)$
  \EndProcedure
\end{algorithmic}
Here $\mathfrak{C}$ is connected-component labeling operator and $\odot$ is
defined as follows:
\begin{equation}
  x \odot y = \left\{
  \begin{array}{ll}
    1 & \quad x = y \\
    0 & \quad \text{otherwise}
  \end{array}
  \right.
\end{equation}

>>>>

\section{Computational algorithms}
Three-point correlation function $S_3$ can be computed either in spatial domain
or in frequency domain via the convolution theorem. Suppose that a finite
countable set $A$ is defined as a multidimensional array. A point $x$ belongs to
a subset $A_n$ if and only if $A[x] = n$. The simplest algorithm below
calculates $S_3$ pointwise in spatial domain:
\begin{algorithmic}[1]
  \Procedure{$S_3$}{$A, n, x_1, x_2$}
  \State $A_n \gets I_n (A)$
  \Comment Apply $I_n$ to the input array elementwise.
  \State $A'_n \gets \mathfrak{S}_{x_1}(A_n)$
  \State $A''_n \gets \mathfrak{S}_{x_2}(A_n)$
  \State $T \gets A_n \cdot A'_n \cdot A''_n$
  \State \textbf{return} $Sum(T) / Norm(A, x_1, x_2)$
  \EndProcedure
\end{algorithmic}
In this algorithm the indicator function $I_n$ for the phase of interest $n$ is
applied to the input array $A$. Then the shift operator $\mathfrak{S}$ is used
to shift $A_n$ by vectors $x_1$ and $x_2$. Finally, the three arrays are
multiplied pointwise and a normalized sum of all elements of the product is
returned.

There are two common boundary conditions when applying $\mathfrak{S}$. The
first condition is to extend $A_n$ periodically (\cref{fig:s3-periodic}) when
accessing out-of-bounds array elements. In this case $\mathfrak{S}_z$ is a
circular shift operator:
\begin{equation}
  A[x] \leftarrow A[x+z \mod d]
\end{equation}
In this case $Norm(A, x_1, x_2)$ does not depend on $x_1$ and $x_2$ and
is equal to the total number of elements in $A$. The second condition is to
replace out-of-bound elements with zeros (so-called zero padding)
(\cref{fig:s3-zeros}). A formula for $Norm(A, x_1, x_2)$ in this case is in
\ref{sec:number-of-trials} and $\mathfrak{S}_z$ is defined as follows:
\begin{equation}
  A[x] \leftarrow \left\{
  \begin{array}{ll}
    A[x+z] & \quad \text{if $x+z$ is a correct index in $A$} \\
    0 & \quad \text{otherwise}
  \end{array}
  \right.
\end{equation}
\begin{figure*}[tp]
  \centering
  \subfigure[Periodic boundary conditions]{
    \includegraphics[width=0.45\linewidth]{images/periodic.png}
    \label{fig:s3-periodic}}
  \hfill
  \subfigure[Zero padding]{
    \includegraphics[width=0.45\linewidth]{images/zeros.png}
    \label{fig:s3-zeros}}
  \caption[]{Computation of unnormalized $S_3$ function at point $(1, 2)$ for
    one-dimensional sequence of length 8.}
  \label{fig:s3-computation}
\end{figure*}

@@

Algorithms for three-point surface functions are the same as for $S_3$,
containing only small modifications. The main difference is that we need to
replace an indicator function $I_n$ with an edge detection operator $E_n$
(see \ref{sec:filter}). Then the algorithms are straightforward:
\begin{algorithmic}[1]
  \Procedure{$F_{sss}$}{$A, n, x_1, x_2$}
  \State $A_n \gets E_n (A)$
  \State $A'_n \gets \mathfrak{S}_{x_1}(A_n)$
  \State $A''_n \gets \mathfrak{S}_{x_2}(A_n)$
  \State $T \gets A_n \cdot A'_n \cdot A''_n$
  \State \textbf{return} $Sum(T) / Norm(A, x_1, x_2)$
  \EndProcedure

  \Procedure{$F_{ssv}$}{$A, n, x_1, x_2$}
  \State $A_n \gets E_n (A)$
  \State $A_{(void)} \gets I_{(void)} (A)$
  \State $A'_n \gets \mathfrak{S}_{x_1}(A_n)$
  \State $A'_{(void)} \gets \mathfrak{S}_{x_2}(A_{(void)})$
  \State $T \gets A_n \cdot A'_n \cdot A'_{(void)}$
  \State \textbf{return} $Sum(T) / Norm(A, x_1, x_2)$
  \EndProcedure

  \Procedure{$F_{svv}$}{$A, n, x_1, x_2$}
  \State $A_n \gets E_n (A)$
  \State $A_{(void)} \gets I_{(void)} (A)$
  \State $A'_{(void)} \gets \mathfrak{S}_{x_1}(A_{(void)})$
  \State $A''_{(void)} \gets \mathfrak{S}_{x_2}(A_{(void)})$
  \State $T \gets A_n \cdot A'_{(void)} \cdot A''_{(void)}$
  \State \textbf{return} $Sum(T) / Norm(A, x_1, x_2)$
  \EndProcedure
\end{algorithmic}

@@

To compute $C_3$ function we need to use a connected-component labeling
algorithm \cite{4728561,PhysRevB.14.3438} to label clusters. Then, again, a
slightly modified algorithm for $S_3$ is used.
\begin{algorithmic}[1]
  \Procedure{$C_3$}{$A, n, x_1, x_2$}
  \State $A_n \gets I_n (A)$
  \State $C_n \gets \mathfrak{C}(A_n)$
  \State $C'_n \gets \mathfrak{S}_{x_1}(C_n)$
  \State $C''_n \gets \mathfrak{S}_{x_2}(C_n)$
  \State $T \gets C_n \odot C'_n \odot C''_n$
  \State \textbf{return} $Sum(T) / Norm(A, x_1, x_2)$
  \EndProcedure
\end{algorithmic}
Here $\mathfrak{C}$ is connected-component labeling operator and $\odot$ is
defined as follows:
\begin{equation}
  x \odot y = \left\{
  \begin{array}{ll}
    1 & \quad x = y \\
    0 & \quad \text{otherwise}
  \end{array}
  \right.
\end{equation}
